The Forecast Trap

### 2022-02-28 manuscript/round3-manuscript.Rmd
:pen:

**A note on models and data**

@@

# Iterative updating: _Sequential Decision Processes_

@@

## Iterative upates and adaptive management

Rather than select a single best model (or best parameter value), a manager could choose to integrate over possible outcomes generated from all candidate models.
Updating posterior distributions over parameters and/or weights assigned to different models are examples of this kind of adaptive management [@Punt2016; @Ludwig1982].
I illustrate the application of such an adaptive management strategy, following classic examples for parameter [@Ludwig1982] or structural [@Smith1981] model uncertainty.
I first consider only the same two models considered in the previous example. 
I later consider a larger suite of 42 models, spanning the parameter space of Gordon-Schaefer curves.
To avoid failure to explore sufficiently, [see exploration-exploitation tradeoff, e.g. @Walters1981],
I assign prior belief of 99% weight on the optimally performing model, model 1.  
For comparison, I consider the baseline case in which the manager does not update posterior distribution over which models/parameter values are correct (the manager still chooses a new TAC after each observation, but does not update their belief in the model, i.e. does not _learn_ over time).
The difference between the performance with and without learning is known as the "Value of Information".
In both 2-model and 42-model scenarios, the value of information is strongly negative.
The 2-model case achieves a net present value to -58% of the value of having used model 1 alone [Fig 4]. 
Including all 42 models reduces this to a value of -32%.
Both harvests and fish biomass remain significantly lower under adaptive learning scenarios.

@@

out <- am %>% group_by(method) %>% summarize(total = sum(quota*.99^time))
( out[out$method=="learning","total"][[1]] -  out[out$method=="planning","total"][[1]]) / out[out$method=="planning","total"][[1]]

out <- am_multi %>% group_by(method) %>% summarize(total = sum(quota*.99^time))
( out[out$method=="learning","total"][[1]] -  out[out$method=="planning","total"][[1]]) / out[out$method=="planning","total"][[1]]

@@

A manager may also escape the forecast trap by exploring actions which are never recommended from any of the available forecasts.
For example, a manager looking at the low stocks and poor harvests achieved under model 2, could decide experimentally to reduce fishing quota. 
This will allow the population to enter a range of state-space where discrepancies between model 2 and observations are more obvious.
@Runge2016 describes how such "double-loop learning" can be used to identify when the entire model set is inaccurate, a problem which is not solved by "single-loop" adaptive-management in our examples above.
@Schindler2015 also underscores the value of flexible policies; rather than "managing solely within the range of past variation; active probing is usually needed," and contrasts this to a typical interpretation of the 'precautionary principle' often cited as a reason to avoid exploratory actions.
However, just because active exploration can escape the forecast trap does mean it is always a good idea.

@@

Finally, it is worth noting that decisions do not need to be premised on a forecast at all, but can be premised entirely on the basis of past experience:
'If the fish stock size has gone up, increase harvest slightly, otherwise, decrease slightly.'
Such a policy is not optimal, but it is robust across a wide range of unimodal stock-recruitment curves without ever estimating a predictive model.
This is the basis of so-called 'model-free' reinforcement learning algorithms such as DQN [@DQN] and SAC [@SAC], which train deep neural networks to learn a policy without ever attempting to predict future states of the underlying process.
Training such a artificial intelligence agents across a wide suite of simulations,
a process known as curriculum learning [@curriculum_learning], mimics the scenario analyses and search for robust policies. Such approaches have been used to train agents to play 2600 Atari console games at superhuman ability [@DQN], outperform race-car drivers [@sony] and control nuclear fusion reactions [@deepmind_fusion].
Such approaches are also not yet well understood, and can fail dramatically [@Henderson2019; @Dulac-Arnold_2019].

>>>>

## A note on models and data

@@

# An Iterative Decision Example

@@

## Adaptive Management

Rather than select a single best model (or best parameter value), a manager could choose to integrate over possible outcomes generated from all candidate models.
Updating posterior distributions over parameters and/or weights assigned to different models are examples of this kind of adaptive management [@Punt2016; @Ludwig1982].
I illustrate the application of such an adaptive management strategy, following classic examples for parameter [@Ludwig1982] or structural [@Smith1981] model uncertainty.
I first consider only the same two models considered in the previous example. 
I later consider a larger suite of 42 models, spanning the parameter space of Gordon-Schaefer curves.
To avoid failure to explore sufficiently, [see exploration-exploitation trade-off, e.g. @Walters1981],
I assign prior belief of 99% weight on the optimally performing model, model 1.  
For comparison, I consider the baseline case in which the manager does not update posterior distribution over which models/parameter values are correct (the manager still chooses a new TAC after each observation, but does not update their belief in the model, i.e. does not _learn_ over time).
The difference between the performance with and without learning is known as the "Value of Information" [@Howard1966].
In both 2-model and 42-model scenarios, the value of information is strongly negative.
The 2-model case achieves a net present value to -58% of the value of having used model 1 alone [Fig 4]. 
Including all 42 models reduces this to a value of -32%.
Both harvests and fish biomass remain significantly lower under adaptive learning scenarios.

@@

out <- am %>% group_by(method) %>% summarize(total = sum(quota*.99^time))
VOI_2models <- ( out[out$method=="learning","total"][[1]] -  out[out$method=="planning","total"][[1]]) / out[out$method=="planning","total"][[1]]

out <- am_multi %>% group_by(method) %>% summarize(total = sum(quota*.99^time))
VOI_42models <- ( out[out$method=="learning","total"][[1]] -  out[out$method=="planning","total"][[1]]) / out[out$method=="planning","total"][[1]]

@@

Second, A manager may also escape the forecast trap by exploring actions which are never recommended from any of the available forecasts.
For example, a manager looking at the low stocks and poor harvests achieved under model 2, could decide experimentally to reduce fishing quota. 
This will allow the population to enter a range of state-space where discrepancies between model 2 and observations are more obvious.
@Runge2016 describes how such "double-loop learning" can be used to identify when the entire model set is inaccurate, a problem which is not solved by "single-loop" adaptive-management in our examples above.
@Schindler2015 also underscores the value of flexible policies; rather than "managing solely within the range of past variation; active probing is usually needed," and contrasts this to a typical interpretation of the 'precautionary principle' often cited as a reason to avoid exploratory actions.
However, just because active exploration can escape the forecast trap does mean it is always a good idea.

@@

Finally, it is worth noting that decisions do not need to be premised on a forecast at all, but can be premised entirely on the basis of past experience:
'If the fish stock size has gone up, increase harvest slightly, otherwise, decrease slightly.'
Such a policy is not optimal, but it is robust across a wide range of unimodal stock-recruitment curves without ever estimating a predictive model.
This is the basis of so-called 'model-free' reinforcement learning algorithms such as DQN [@DQN] and SAC [@SAC], which train deep neural networks to learn a policy without ever attempting to predict future states of the underlying process.
Training such a artificial intelligence agents across a wide suite of simulations,
a process known as curriculum learning [@curriculum_learning], mimics the scenario analyses and search for robust policies. Such approaches have been used to train agents to play 2600 Atari console games at superhuman ability [@DQN], outperform race-car drivers [@sony] and control nuclear fusion reactions [@deepmind_fusion].
Such approaches are also not yet well understood, introducing new risks as well as new possibilities[@Henderson2019; @Dulac-Arnold_2019].
